### 2019-04-07 papers/journal/experiments.tex
Minor update
Dia is complet
Section 7.1 completed

%\externaldocument{sat-table}
%We have developed four different tools (\sattool, \smttool, \qbftool, \ourtool) for implementation of proposed encodings.
%
\noindent We have developed the {\vtstool} tool to conjecture the properties of the VTS (explained in Section 6). 
%
%
The {\vtstool} tool includes about 10,000 lines of code.
%
The tool is divided into three different blocks,
%
checking existential condition, checking the universal condition and performing synthesis on the incomplete VTS. 
%
The existential condition is subdivided into two sub-block, bounded model checking and constraint solving. 
%
We will refer to existential subblocks as {\sattool} and {\smttool}, universal block as {\qbftool} and synthesis block as {\ourtool}.
%

\subsection{Implementation}
%The existential condition problem is implemented with two separate encodings.
\noindent We have implemented two separate encodings to determine the existential condition of the VTS. 
%To solve the existential condition we have implemented two separate encodings. 
%
The first encoding is implemented in {\sattool}
%One 
%with a successful bounded model checker, CBMC, 
and perform path enumeration to encode the reachability. 
%
%This encoding is implemented in {\sattool}.
% and uses a successful bounded model checker, CBMC.
%which is based on the architecture of a
%
%The second encoding is implemented in {\smttool} and uses a
%is based on 
%industrial strength SMT solver, Z3, that uses a 
%recursive encoding of finding a least fixed point to encode the same. 
The second encoding is implemented in {\smttool} and employs a recursive encoding of finding a least fixed point for the reachability.

%
%The encoding is implemented in {\smttool}.
% and uses industrial strength SMT solver, Z3.
%The {\sattool} is based on the architecture of a bounded model checker (CBMC) and uses an enumeration of paths for encoding the reachability problem.
%
%The {\smttool} is based on the SMT solver Z3 and uses the efficient encoding of finding a least fixed point to solve reachability problem. 
%
%The {\qbftool} is based on the QBF encoding of the universal condition and uses {\depqbf} solver to solve the generated QBF. 
The QBF encoding of the universal condition is implemented in {\qbftool}.
%
The tool uses {\depqbf} solver to solve the generated QBF formula. 
%
The {\ourtool} is a synthesis tool for synthesizing incomplete information in the input VTS using {\qbftool} interface.
%

@@

\textbf{{\sattool (Bounded model checker based tool)}}
The $\fbmc$ encoding based on enumeration of paths (presented in section 6.2) for each variants of VTS is implemented in {\sattool}.
	%We have implemented the encodings for different variants of VTS (presented in section 6.2) using
The tool is based on CBMC, a successful bounded model checker, for C and C++ programs. 
%
%Unlike a general model checker~\cite{cimatti2002nusmv}, CBMC
%
It verify property for all the possible states, up to a certain “depth,” a parametric limit on the size of the model.
%
The depth of model is specified by the user and can be  incrementally increased to a desired value.
%
%The CBMC architecture consists of a C-language front-end (CFE) and a SAT Solver back-end (SAT). 
%
%CBMC front end accepts

%We encode the problem using CBMC C-language front-end (CFE) interface with some special annotations called \textit{assumes} and \textit{assertions}.
%
%These annotations express constraints on the model and properties to be checked. 
%	
The vesicle transport network was modelled as a non-deterministic C-program manipulating an annotated graph. 
%
The special annotations \textit{assumes} and \textit{assertions} of CBMC's front end were used to express constraints on the model and properties to be checked. 
%
%The fusion rules and steady state properties were expressed as constraints using special annotations \textit{assumes}. 
%		
%The correspondence relationship between connectedness conditions and steady state were expressed as annotations \textit{assertions} to be checked. 
Using the tool's interface user can select the variation of the VTS, connectivity of the graph and other required parameters like an exploration depth.
%
The {\sattool} verifies the properties by first converting the constraints and properties into a Boolean formula and determine the satisfiability of the formula with the help of a SAT solver backend.
%
The satisfying model is translated into a human readable format and in case of a violation, the tool produces a readable counterexample.
%  following steps:
%	\begin{itemize}
%	\item Convert (using CFE) the model and the properties into a Boolean formula (verification condition) such that the property is true of all behaviors of the model up to the specified
%depth iff the Boolean formula is valid.
%	\item Check (using a SAT solver) validity of the verification condition. {\sattool} reports successful verification if the formula is found to be valid. If not, then {\sattool} produces a counterexample,
%an assignment to the variables of the model, that is a witness to the violation of the property.
%	\end{itemize}
For our experimentation purpose, the SAT solver backend used were MiniSat~\cite{sorensson2005minisat}, and CryptoMiniSat~\cite{soos2016cryptominisat}.
% the winner of SAT 2015 Competition~\cite{balyo2016sat}.

@@

\textbf{{\smttool} (SMT based tool)}
The $\fsmt$ encoding based on recursive reachability definition (presented in section 6.1)  
%We have implemented the encodings 
for each variant is implemented in {\smttool}.
%
The tool is based on industrial strength SMT solver, Z3. 
%
The tool allows the user to choose a model and the size
of the network besides other parameters like connectivity and number of parallel edges. 
%		
It uses Z3 Python interface to build the needed constraints and applies Z3 solver on the constraints to find a model (a satisfying assignment that follows the constraints). 
%
This tool also translates the satisfying model found by Z3 into
a VTS and represent a visual output to the user in form of coloured annotated graph. 
%
%We also visually report the dropped edges required to disconnect the graph, it gives information about the connectivity of the graph.
%

\textbf{{\qbftool} (QBF based universal condition)}
The $\fqbf$ encoding (presented in section 6.5) is implemented in the {\qbftool}. 
%
The tool is developed in C++ and uses~\zthree~\cite{z3} infrastructure to encode the constraints.
%
Since some of the formulas involve alternation of quantifiers over Boolean variables Z3 is not a suitable choice for those examples.
%
We translate the formulas created by Z3 tool into a standard
QDIMACS~\cite{qdimacs} format and use QBF solvers to solve the formula. 
%
We have used~\depqbf~\cite{lonsing2010depqbf} for solving QBF formulas with alternation of quantifiers.

\textbf{{\ourtool} (QBF based synthesis tool)}
\input{synth-table}
The synthesis encodings (presented in section 6.6) is implemented  in~\ourtool\footnote{{\url{https://github.com/arey0pushpa/pyZ3}}}.
%
The tool takes a partially defined VTS as input in a custom designed
input language.
%
The input is then converted to the constraints over VTS using the {\qbftool} interface. 
%
The tool can synthesize the queries discussed earlier and also their combinations.
%The tool can not only synthesize the discussed queries, but also their combinations.
%
For example, our tool can modify labels of nodes or edges while
learning activity functions.
%
The tool returns a complete VTS with visual graph representation as an output to the user.

@@

%
Table~\ref{tab:satsmt-graph} presents the comparison of the running times of {\smttool} and {\sattool} on VTS variations of sizes 2 to 10.
%
For this comparison we have fixed the total number of molecules minimum four times the node count 
%for $ |N|> 2$ and $2|N| + 1$ for $|N| = 2$
%
%For each variant, we fix 1
and maximum number of parallel
edges to be 2.
%
For our analysis we relax the 3:1 requirement for the Q and R-SNARES and treat a Q-SNARE triple helix as a single complex as modelled in~\cite{shukla2017discovering}.
%
The 
%
In the table M/O indicates the search results in out of memory, T/O indicates a time out and ``!'' indicate that no model exists for the constraint  (constraints were unsatisfiable).
%In the table we have shown comparison for specific connectivity, for example variant A is checked against any graph with connectivity 2, variant B with connectivity 3 similarly for the rest of the variants.
%
%We have compared our performance with the performance of our earlier CBMC based implementation (old-encoding).
%
As an example, the result of verifying the formula for variation C of 2-connected, 10 compartment graph took 1977.72 secs for a satisfying result.
%
In comparison, {\sattool} was unable to solve the instance and ran out of memory.
% for $|N|$ greater than 5.
%
% 
The recursive encoding also performed better for the cases where the instance was unsatisfiable (Variant A timing comparison).
%instead of enumeration, not only we got efficiency improved for finding a SAT model but also did better in the case of refutation that no model exist (table~\ref{tab:satsmt-graph} Variant A timing comparison and Variant D with N =2). 
%
Hence with the use of improved encoding of reachability, we were able to scale the system to the biologically relevant compartment size, the
%a much larger compartmentalized system, especially to interesting 
the case of eukaryotic cells with a total number of 10 compartments.
%
Furthermore, we experimented with limits of our tools and found that {\smttool} was able to solve the constraints up to $18$ nodes.

>>>>

%\externaldocument{sat-table}
%We have developed four different tools (\sattool, \smttool, \qbftool, \ourtool) for implementation of proposed encodings.
%
\noindent We have developed the {\vtstool} tool to conjecture the properties of the VTS (explained in Section 6). 
%
%
The {\vtstool} tool includes about 10,000 lines of code.
%
The tool is divided into three different blocks,
%
checking existential condition, checking the universal condition and performing synthesis on the incomplete VTS. 
%
The existential condition is subdivided into two sub-blocks, bounded model checking and constraint solving. 
%
We will refer to existential subblocks \unsure{The fonts of toolset looks different make them uniform. The reason is valid although.}as {\sattool} and {\smttool}, universal block as {\qbftool} and synthesis block as {\ourtool}.
%

\subsection{Implementation}
%The existential condition problem is implemented with two separate encodings.
\noindent We have implemented two separate encodings to determine the existential condition of the VTS. 
%To solve the existential condition we have implemented two separate encodings. 
%
The first encoding is implemented in {\sattool}
%One 
%with a successful bounded model checker, CBMC, 
and perform path enumeration to encode the reachability. 
%
%This encoding is implemented in {\sattool}.
% and uses a successful bounded model checker, CBMC.
%which is based on the architecture of a
%
%The second encoding is implemented in {\smttool} and uses a
%is based on 
%industrial strength SMT solver, Z3, that uses a 
%recursive encoding of finding a least fixed point to encode the same. 
The second encoding is implemented in {\smttool} and employs a recursive encoding
% of finding a least fixed point 
for the reachability. 
%
%The encoding is implemented in {\smttool}.
% and uses industrial strength SMT solver, Z3.
%The {\sattool} is based on the architecture of a bounded model checker (CBMC) and uses an enumeration of paths for encoding the reachability problem.
%
%The {\smttool} is based on the SMT solver Z3 and uses the efficient encoding of finding a least fixed point to solve reachability problem. 
%
%The {\qbftool} is based on the QBF encoding of the universal condition and uses {\depqbf} solver to solve the generated QBF. 
The QBF encoding of the universal condition is implemented in {\qbftool}.
%
The tool uses {\depqbf} solver to solve the generated QBF formula. 
%
The {\ourtool} is a synthesis tool for synthesizing incomplete information in the input VTS using {\qbftool} interface.
%

@@

%\textbf{{
\subsubsection{{\sattool} (Bounded model checker based tool)}
%}}
The $\fbmc$ encoding based on enumeration of paths (Section~\ref{enc:bmc}) for each variants of VTS is implemented in {\sattool}.
	%We have implemented the encodings for different variants of VTS (presented in section 6.2) using
The tool is based on CBMC, a successful bounded model checker, for C and C++ programs. 
%
%Unlike a general model checker~\cite{cimatti2002nusmv}, CBMC
%
It verifies property for all the possible states, up to a certain “depth,” a parametric limit on the size of the model.
%
The depth of the model is specified by the user and can be incrementally increased to the desired value.
%
%The CBMC architecture consists of a C-language front-end (CFE) and a SAT Solver back-end (SAT). 
%
%CBMC front end accepts

%We encode the problem using CBMC C-language front-end (CFE) interface with some special annotations called \textit{assumes} and \textit{assertions}.
%
%These annotations express constraints on the model and properties to be checked. 
%	
The vesicle transport network was modeled as a non-deterministic C-program manipulating an annotated graph. 
%
The special annotations \textit{assumes} and \textit{assertions} of CBMC's front end were used to express constraints on the model and properties to be checked. 
%
%The fusion rules and steady state properties were expressed as constraints using special annotations \textit{assumes}. 
%		
%The correspondence relationship between connectedness conditions and steady state were expressed as annotations \textit{assertions} to be checked. 
Using the tool's interface user can select the variation of the VTS, connectivity of the graph and other required parameters like an exploration depth.
%
The {\sattool} verifies the properties by first converting the constraints and properties into a Boolean formula and determine the satisfiability of the formula with the help of a SAT solver backend.
%
The satisfying model is translated into a human readable format and in case of a violation, the tool produces a readable counterexample.
%  following steps:
%	\begin{itemize}
%	\item Convert (using CFE) the model and the properties into a Boolean formula (verification condition) such that the property is true of all behaviors of the model up to the specified
%depth iff the Boolean formula is valid.
%	\item Check (using a SAT solver) validity of the verification condition. {\sattool} reports successful verification if the formula is found to be valid. If not, then {\sattool} produces a counterexample,
%an assignment to the variables of the model, that is a witness to the violation of the property.
%	\end{itemize}
For our experimentation purpose, the SAT solver backend used were MiniSat~\cite{sorensson2005minisat} and CryptoMiniSat~\cite{soos2016cryptominisat}.
% the winner of SAT 2015 Competition~\cite{balyo2016sat}.

@@

\subsubsection{{\smttool} (SMT based tool)}
The $\fsmt$ encoding based on recursive reachability definition (Section~\ref{enc:smt})  for each variant is implemented in {\smttool}.
%
The tool is based on industrial strength SMT solver, Z3. 
%
The tool allows the user to choose a model and the size
of the network besides other parameters like connectivity and number of parallel edges. 
%        
It uses the Z3 Python interface to build the needed constraints and applies Z3 solver on the constraints to find a model (a satisfying assignment that follows the constraints). 
%
This tool also translates the satisfying model found by Z3 into
a VTS and represent a visual output to the user in the form of a colored annotated graph. 
%
%We also visually report the dropped edges required to disconnect the graph, it gives information about the connectivity of the graph.
%

\subsubsection{{\qbftool} (QBF based universal condition)}
The $\fqbf$ encoding (Section~\ref{enc:univ-cond}) is implemented in the {\qbftool}. 
%
The tool is developed in C++ and uses~\zthree~\cite{z3} infrastructure to encode the constraints.
%
Since some of the formulas involve alternation of quantifiers over Boolean variables Z3 is not a suitable choice for those examples.
%
We translate the formulas created by Z3 tool into a standard
QDIMACS~\cite{qdimacs} format and use QBF solvers to solve the formula. 
%
We have used~\depqbf~\cite{lonsing2010depqbf} for solving QBF formulas with alternation of quantifiers.

\subsubsection{{\ourtool} (QBF based synthesis tool)}
\input{synth-table}
The synthesis encodings (Section~\ref{enc:synth}) is implemented  in~\ourtool\footnote{{\url{https://github.com/arey0pushpa/pyZ3}}}.
%
The tool takes a partially defined VTS as input in a custom designed
input language.
%
The input is then converted to the constraints over VTS using the {\qbftool} interface. 
%
The tool can synthesize the queries discussed earlier and also their combinations.
%The tool can not only synthesize the discussed queries, but also their combinations.
%
For example, our tool can modify labels of nodes or edges while
learning activity functions.
%
The tool returns a complete VTS with visual graph representation as an output to the user.

@@

%
Table~\ref{tab:satsmt-graph} presents the comparison of the running times of {\smttool} and {\sattool} on VTS variations of sizes 2 to 10.
%
For this comparison we have fixed the total number of molecules minimum four times the node count 
%for $ |N|> 2$ and $2|N| + 1$ for $|N| = 2$
%
%For each variant, we fix 1
and maximum number of parallel
edges to be 2.
%
%For our analysis we relax the 3:1 requirement for the Q and R-SNARES and treat a Q-SNARE triple helix as a single complex as modelled in~\cite{shukla2017discovering}.
%
%The 
%
In the table M/O indicates the search results in out of memory, T/O indicates a time out and ``!'' indicate that no model exists for the constraint  (constraints were unsatisfiable).
%In the table we have shown comparison for specific connectivity, for example variant A is checked against any graph with connectivity 2, variant B with connectivity 3 similarly for the rest of the variants.
%
%We have compared our performance with the performance of our earlier CBMC based implementation (old-encoding).
%
As an example, the result of verifying the formula for variation C of 2-connected, 10 compartment graph took 1977.72 secs for a satisfying result.
%
In comparison, {\sattool} was unable to solve the instance and ran out of memory.
% for $|N|$ greater than 5.
%
% 
The recursive encoding also performed better for the cases where the instance was unsatisfiable (Variant A timing comparison).
%instead of enumeration, not only we got efficiency improved for finding a SAT model but also did better in the case of refutation that no model exist (table~\ref{tab:satsmt-graph} Variant A timing comparison and Variant D with N =2). 
%
Hence with the use of improved encoding of reachability, we were able to scale the system to the biologically relevant compartment size, the
%a much larger compartmentalized system, especially to interesting 
the case of eukaryotic cells with a total number of 10 compartments.
%
Furthermore, we experimented with limits of our tools and found that {\smttool} was able to solve the constraints up to $18$ nodes.
